Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-12 emsoft-2019/Motiv.tex
fixed example description

As a smaller example, consider the standard problem of designing a controller for an inverted pendulum depicted in~\autoref{fig:inverted_pendulum}~{(a)}.
The goal of the controller is to keep the pendulum at the vertical position while satisfying hard constraints on the state variables and control inputs.
A model of the system can be constructed using physical principles. 
After linearization and time discretization, the model is
	\begin{equation}
		\begin{bmatrix}
			 \theta_{k+1}\\
			\omega_{k+1}
		\end{bmatrix}=
		\begin{bmatrix}
			1 & T_s\\
			\frac{T_sg}{L}& (1-\frac{T_sb}{mL^2})		
		\end{bmatrix}
		\begin{bmatrix}
			\theta_k\\
			\omega_k
		\end{bmatrix}+
		\begin{bmatrix}
			0\\
			\frac{T_s}{mL^2}
		\end{bmatrix}u_k + w_k
		\label{eq:pendul_ss}
	\end{equation}
where $\theta_k$, $\omega_k$, and $u_k$ denote respectively the angular position, angular speed, 
and the input torque at time $kT_s$ with $T_s$ being an appropriate sample time.
The disturbance $w_k\in\mathbb R^2$, bounded with a polyhedral set $w_k\in \mathcal W$, captures the modeling error due to 
linearization and discretization. 
The parameter $g=9.81 [m/s^2]$ is the gravitational acceleration, $m$ is the ball mass, $b$ is the rotational fraction coefficient, 
and $L$ is the length of the bar. 
Starting from an initial state $(\theta_0,\omega_0)$, the control goal is to converge to the equilibrium point 
$\theta=0, \omega=0$.
Additionally, we require the state constraints $\theta_k\in[-\pi,\pi]$ and $\omega_k\in[-\pi/8,\pi/8]$ to hold at all time instances.

@@

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use MATLAB to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
\RM{the following setence is not understandable. This is the first time we introduce incorrect region selection.
Who is Daisy?}
\RM{A better version is to informally define incorrect region selection. Say that we run an algorithm for incorrect region
selection to bound the possible error. 
Break the sentence into two. Introduce Daisy as a mixed-precision tuning tool.}
Over the computed controller, we compute the error due to an incorrect region selection $e$ and in case that $e<\Delta_0$, we ask Daisy
to provide a mixed-precision scheme such that the associated implementation error is at most $\Delta_0$, that is,
at all points in the control domain, the difference between the infinite-precision controller function
and its mixed-precision implementation is at most $\Delta_0$.

Unfortunately, $\Delta_0=0.01$ is not realizable, as the region selection error is $0.019$ and hence $e>\Delta_0$. 
%always greater than zero either due to selecting the wrong region in the affine controller function or due to the finite-precision computation of the control action.
%\eva{Is this the error due to mixed-precision or due to wrong regions?}
Therefore, we define $\Delta_1 := \Delta_0 + \varepsilon_{SAFE}$ with $\varepsilon_{SAFE}$ being a fixed threshold (we chose $\varepsilon_{SAFE} = 0.04$) and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller,
shown in Figure~\ref{fig:inverted_pendulum}(b).
The controller has $14$ regions. Following demostrates one sample mapping, including the activation condition and its corresponding affine function
\begin{equation*}
\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02$}
\end{equation*}
where, output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. Notice that all the rest of mappings have similar format.
%\RM{give the control function. I think it will be good to explicitly write the polyhedra and the affine maps}

@@

Next, we ask Daisy to provide us with mixed-precision implementation for all the parameters and to compute a static bound on the
implementation error. 
The mixed-precision scheme given by Daisy results in an implementation with $6084$ bits in total 
for the corresponding robust explicit MPC.
Daisy also computes a static error bound of $0.03$.
Since this error bound is less than our robustness margin $\Delta_1 = 0.05$, we terminate the loop
and return this implementation.
The mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.

>>>>

As a smaller example, consider the standard problem of designing a controller for an inverted pendulum depicted in~\autoref{fig:inverted_pendulum}~{(a)}.
The goal of the controller is to keep the pendulum at the vertical position while satisfying hard constraints on the state variables and control inputs.
A model of the system can be constructed using physical principles. 
After linearization and time discretization, the model is
	\begin{equation}
		\begin{bmatrix}
			 \theta_{k+1}\\
			\omega_{k+1}
		\end{bmatrix}=
		\begin{bmatrix}
			1 & T_s\\
			\frac{T_sg}{L}& (1-\frac{T_sb}{mL^2})		
		\end{bmatrix}
		\begin{bmatrix}
			\theta_k\\
			\omega_k
		\end{bmatrix}+
		\begin{bmatrix}
			0\\
			\frac{T_s}{mL^2}
		\end{bmatrix}u_k + w_k
		\label{eq:pendul_ss}
	\end{equation}
where $\theta_k$, $\omega_k$, and $u_k$ denote respectively the angular position, angular speed, 
and the input torque at time $kT_s$ with $T_s$ being an appropriate sampling time.
The disturbance $w_k\in\mathbb R^2$, bounded with a polyhedral set $w_k\in \mathcal W$, captures the modeling error due to 
linearization and discretization. 
The parameter $g=9.81 [m/s^2]$ is the gravitational acceleration, $m$ is the ball mass, $b$ is the rotational fraction coefficient, 
and $L$ is the length of the bar. 
Starting from an initial state $(\theta_0,\omega_0)$, the control goal is to converge to the equilibrium point 
$\theta=0, \omega=0$.
Additionally, we require the state constraints $\theta_k\in[-\pi,\pi]$ and $\omega_k\in[-\pi/8,\pi/8]$ to hold at all time instances.

@@

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use Matlab to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
% \RM{the following setence is not understandable. This is the first time we introduce incorrect region selection.
% Who is Daisy?}
% \RM{A better version is to informally define incorrect region selection. Say that we run an algorithm for incorrect region
% selection to bound the possible error. 
% Break the sentence into two. Introduce Daisy as a mixed-precision tuning tool.}

The output of robust EMPC given by Matlab decomposes the control domain into a
finite number of polyhedral domains together with an affine map for each domain. At each
time step, the state vector $x_k$ is read (or estimated) and based on the domain
that it belongs to, the corresponding control output is computed. 
However, the polyhedral regions are implemented in finite precision so that
it is possible that due to quantization errors the domain is selected incorrectly.
We refer to the resulting error as incorrect region selection error.
% However, due to error by fixed-point implementation, it is possible that the domain is selected
% incorrectly. This causes an error term which we refer to it as incorrect region
% selection error.

In our example, we compute an incorrect region selection error $e = 0.019$, which is larger
than $\Delta_0=0.01$. We therefore increase the disturbance bound by a fixed amount to
$\Delta_1 = 0.05$ and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
%
% Over the high-precision controller, we compute the error due to an incorrect region selection $e$. In case that $e<\Delta_0$, we would like to find the smallest number of bits required for implementing the controller such that we are still within the safe bound $\Delta_0$. To do so, we ask Daisy, which is a mixed-precision tuning tool, to provide us with a mixed-precision scheme such that the associated implementation error is at most $\Delta_0$, that is,
% at all points in the control domain, the difference between the infinite-precision controller function
% and its mixed-precision implementation is at most $\Delta_0$.
%
%Unfortunately, $\Delta_0=0.01$ is not realizable, as the region selection error is $0.019$ and hence $e>\Delta_0$. 
%always greater than zero either due to selecting the wrong region in the affine controller function or due to the finite-precision computation of the control action.
%\eva{Is this the error due to mixed-precision or due to wrong regions?}
% Therefore, we define $\Delta_1 := \Delta_0 + \varepsilon_{SAFE}$ with $\varepsilon_{SAFE}$ being a fixed threshold (we chose $\varepsilon_{SAFE} = 0.04$) and enlarge the disturbance set as $\mathcal W_{\Delta_1}$. 
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller
with 14 regions, shown in~\autoref{fig:inverted_pendulum}(b).
A sample region with the corresponding affine map is shown below:
%Following demostrates one sample mapping, including the activation condition and its corresponding affine function
\begin{equation*}
\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02$}
\end{equation*}
where the output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. 
The remaining mappings have a similar structure.
For this new controller, the incorrect region selection error is still
approximately $e = 0.019$, which is now below the disturbance bound $\Delta_1$,
so that we can continue with the precision assignment.

@@

Next, we use the precision tuning tool Daisy to provide us with a mixed-precision 
implementation for all the parameters which satisfies the error bound $0.031$,
which is given by the remaining error budget $\Delta_1 - e = 0.05 - 0.019$.  
The mixed-precision implementation returned by Daisy requires $6084$ bits in total 
for the corresponding robust explicit MPC.
% Since this error bound is less than our robustness margin $\Delta_1 = 0.05$, we terminate the loop
% and return this implementation.
The mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.
